---
title: "index"
author: "Alison Haupt"
date: "11/30/2017"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## `geom_point`  

```{r Scatterplot}

#mpg

## plot mpg
g <- ggplot(data = mpg, aes(x = displ, y = hwy))
g + geom_point()
```

### `color`!

```{r addSize}
g + geom_point(aes(color = class))

# could also do
ggplot(data = mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point()

```

### change `color` to `shape`
```{r BWplot}
ggplot(data = mpg, aes(x = displ, y = hwy, shape = class)) +
  geom_point()
```

### change `color` to `size`
```{r size}
ggplot(data = mpg, aes(x = cty, y = hwy, size = class, color = fl)) +
  geom_point(alpha = 0.5)
```
### fix `color` 
```{r fix color}
ggplot(data = mpg, aes(x = cty, y = hwy, size = class)) +
  geom_point(alpha = 0.5, color = "blue")
```
### Exercises
What’s gone wrong with this code?
    ggplot(data = mpg) + 
      geom_point(aes(x = displ, y = hwy, color = "blue"))
      
```{r FixCode}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
     geom_point(color = "blue")
```

```{r lessthan}

ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  aes(color = displ < 5)
     geom_point()
```
## other geoms
```{r other}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
     geom_jitter()

ggplot(data = mpg, aes(x = drv, y = hwy)) + 
     geom_boxplot()

ggplot(data = mpg, aes(x = drv, y = hwy)) + 
     geom_violin()
```
## add geoms on top
```{r other2}
ggplot(data = mpg) + 
    geom_point(aes(x = displ, y = hwy)) +
    geom_smooth(aes(x = displ, y = hwy))

# or could move aes into ggplot
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
    geom_point() +
    geom_smooth()
```
### customize plot
```{r customize}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
    geom_point(aes(color = class)) +
    geom_smooth(size = 1) +
    labs(title = "Relationship between engine size and MPG", x = "Highway MPG", y = "Engine displacement (liters)") + 
    theme_classic() +
    theme(text = element_text(size = 16)) +
    scale_color_discrete(name = "color = class") 
    

ggplot(data = mpg, aes(x = displ, y = hwy)) + 
    geom_point() +
    facet_wrap(~ class) +
    geom_smooth(size = 3) +
    labs(title = "Relationship between engine size and MPG", x = "Highway MPG", y = "Engine displacement (liters)") + 
    theme_classic() +
    theme(text = element_text(size = 16))


```

## Data wrangling with `dplyr`

`read_csv` is from the `readr` package (part of tidyverse). it is not `read.csv`, which is part of base R.
```{r gapminder}

gapminder <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")

## useful functions for getting to know data
head(gapminder)
tail(gapminder)
head(gapminder, 30)
str(gapminder)
names(gapminder)
dim(gapminder)
ncol(gapminder)
nrow(gapminder)

# make dim from ncol and nrow
c(nrow(gapminder), ncol(gapminder))

summary(gapminder)

# Everything above, we were operating on entire data set.  Can also look at just one column

length(unique(gapminder$country))

```

## Actually working with dplyr
```{r dplyr}

filter(gapminder, lifeExp <29)

filter(gapminder, country == "Mexico")

Sweden <- (filter(gapminder, country == "Sweden"))
mean(Sweden$lifeExp)

## or with pipe operator %>% keyboard short cut control shift m  
gapminder %>%  head()
# can still ask for just 3 lines
gapminder %>%  head(3)

filter(gapminder, country == "Sweden") %>% summarize(mean(lifeExp))

# or cleaner

y <- gapminder %>% 
  filter(country == "Sweden") %>% 
  summarize(mean_lifeExpmean = mean(lifeExp))



```

Everything above, we were operating on entire data set.  Can also look at just one c

